\chapter{Visual Analytics of Spatiotemporal Social Media for Abnormal Event Detection}

%Many social media services have emerged in the past years 
%and they provide different types of information according to the purpose of the services.
Social media services, e.g, Twitter, Youtube, Flickr, provide a rich and freely accessible 
database of user-generated situation reports. As advances in technology have enabled
%and the user information tends to contain mainly 
%user's spatial and temporal information while the social media content is being generated.
%Since smart mobile devices became popular recently,
%it is even easier to produce lots of social media data anywhere and anytime. 
%mobility of the 
the widespread adoption of GPS enabled mobile communication
devices, these reports are able to capture important local events observed by an active and ubiquitous community.
The different forms of social media content provided by the users,
such as microposts, images or video footage, can have immense value 
for increasing the situational awareness of ongoing events.

% However, as data volumes have increased beyond the capabilities of manual evaluation, 
% there is a need for advanced tools %plays an important role to provide
% as a means to provide insights for investigation and to understand the extent, 
% severity and consequences of incidents, as well as their time-evolving nature.
% Due to the large number of individual social media messages it is not straightforward 
% to analyze and extract meaningful information. 
% For example, in Twitter, more than 200 million Tweets are posted each day~\cite{Tweets:2011:Twitter}.
% %and a billion Tweets are created
% %every five days as of June 30, 2011.
% Thus, in an ongoing scenario, the relevant messages for situational awareness are usually buried by
% a majority of irrelevant data. 
% Finding and examining these messages without smart aggregation, automated text analysis and 
% advanced filter strategies is almost impossible and extracting meaningful information is even more challenging.

However, as data volumes have increased beyond the capabilities of manual evaluation, there is a need for
advanced tools to aid understanding of the extent, severity and consequences of incidents, 
as well as their time-evolving nature, and to aid in gleaning investigative insights. 
Due to the large number of individual social media messages it is not straightforward to analyze and extract meaningful information. 
For example, in Twitter, more than 200 million Tweets are posted each day~\cite{Tweets:2011:Twitter}. 
Thus, in a developing event, the relevant messages for situational awareness are usually buried by a majority of irrelevant data. Finding and examining these messages
without smart aggregation, automated text analysis and advanced filtering strategies is almost impossible and
extracting meaningful information is even more challenging.

%Another issue is reliability of the data.
%Posts might be outdated, incorrect, or simply unimportant
%and this can heavily interfere with searching for meaningful information. 
%Previously researchers have focused
%on finding spatial and temporal trends according to volume-based importance.
%In other words, a large volume of keywords or content is considered to be more important.
%This type of analysis can cause relatively small volumes of important data to be easily obscured.
%by huge amount of data
%which tends to indicate normal situations, but
%abnormality within a small volume of data can provide more critical and important information for unusual events.

To address these challenges, we present an interactive spatiotemporal social media analytics approach
for abnormal topic detection and event examination~\cite{CHAE:2012:SSM}.
In order to find relevant information within a user defined spatiotemporal frame
%within the social media data, 
we utilize the LDA topic model~\cite{Blei:2003:LDA}, which extracts 
and probabilistically ranks major topics contained in textual parts of the social media data.
%The abnormality is estimated from the major topical events.
The ranks of the categorized topics generally provide a volume-based importance,
but this importance does not reflect the abnormality or criticality of the topic. 
In order to obtain a ranking suitable for situational awareness tasks, we discard daily chatter
by employing the STL~\cite{Cleveland:1990:SAS}.
%based on the abnormality estimation.
In our work, globally and seasonally trending portions of the data are considered less important,
whereas major non-seasonal elements are considered anomalous and, therefore, relevant.

However, due to the large volumes of data, the very specific syntax and semantics of microposts and 
the complex needs of situational analysis, it would not be feasible to apply these techniques 
in the form of a fully automated system. 
Therefore, our whole analysis process, including the application of automated tools, is guided and 
informed by an analyst using a highly interactive visual analytics environment. 
It provides tight integration of semi-automated text-analysis and probabilistic event detection tools 
together with traditional zooming, filtering and exploration following the Information-Seeking Mantra~\cite{Shneiderman:1996:TEH}.

%We apply our techniques to various social media data sources, including Twitter, Flickr, and YouTube,
%to improve reliability of the 
%as a way of testing our abnormality detection scheme. 
%Moreover, it is possible to compare and analyze which information users post according to
%the social media services.
%Finally, our geovisual analytics system is used for situational awareness of
%abnormality within social media data and enables a scalable
%visual examination and highly interactive visual analysis workbench for spatiotemporal trend analysis.
%interactive capability of spatiotemporal trend analysis. 
%The remainder of this document is structured as follows: 
%Section~\ref{sec:relatedwork} is a review of related work.
%The automated methods to find and examine unusual topics and events are described in Section~\ref{sec:social_analytics}.
%In Section~\ref{sec:analysisprocess} we briefly introduce our visual analytics system Scatterblogs, which was already featured in previous works, and explain how the automated methods are integrated within a sophisticated iterative analysis loop.
%Finally we demonstrate the performance of our system based on selected case studies in Section~\ref{sec:case_study} and discuss the approach in Section~\ref{sec:disc}.

\input{ad_process}
\input{ad_LDA}
\input{ad_STL}
\input{ad_analysis}
\input{ad_casestudy}
\input{ad_conclusion}

